Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-20 paper.qmd
Updated supplementary material section

With the model fit, how the NLDR methods organize points in \gD{} layout is examined using the simulated $4\text{-}D$ data having five Gaussian clusters. tSNE, UMAP, and PaCMAP preserved the global structure having well-separated clusters, while tSNE, and UMAP effectively maintains the local structure, showing the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions. The interesting result is that the model fit of PaCMAP show *pancake* shape.

@@

Appendix: The appendix includes more details about the hexagonal binning algorithm (appendix.pdf, Portable Document Format file).

>>>>

With the model fit, some quirks can be observed. The simulated $4\text{-}D$ data having five Gaussian clusters is used. tSNE, and UMAP which showed the fitted model as *crumpled \gD{} sheet* that stretches in all four dimensions evident of preserving local structure. The flat model generated by PaCMAP evident the fail to capture the within cluster variation.

@@

Appendix: The appendix includes more details about the hexagonal binning algorithm and scDEED method (appendix.pdf, Portable Document Format file).
